---
title: "Final Project"
author: "Sashi R Amatya"
date: "2024-11-14"
output:
  word_document: default
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing libraries
```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(data.table)
library(plotly)
```

# Reading excel file
# and then converting Sleep duration to decimal hour format so that 
# 9 hour 30 minutes will show as 9.5 hrs for analysis purpose
```{r}
sleep_data <- read_excel("~/Library/CloudStorage/OneDrive-Personal/SASHI - DATA 211/FINAL PROJECT - SLEEP/Sleep Data - Hours slept.xlsx")
head(sleep_data,3)

# Converting "Sleep Duration" to Hours and Minutes format.
# 9 hours 30 minutes with date will save as = 930
sleep_data$`Sleep Duration`<-
  as.numeric(format(as.POSIXct(sleep_data$`Sleep Duration`),format = "%H%M", units="minutes") )
print(sleep_data)
# Converting "Sleep Duration" minutes to percentage of minute format
# 930 (9 hours 30 minutes) from previous step will now convert to ~9.5 hours 
for(i in 1:nrow(sleep_data)){
  sleep_data$`Sleep Duration`[i] <- (sleep_data$`Sleep Duration`[i]%/%100) + 
          ((sleep_data$`Sleep Duration`[i]%%100)*0.0166)
}
print(sleep_data)
# class(sleep_data$`Sleep Duration`)
```


# Calculating total average sleep per person
```{r}

# Average sleep per person total
avg_sleep <- sleep_data %>%
  group_by(Participant) %>%
  summarise(AverageSleep = mean(`Sleep Duration`)) 

# Graphing - Average sleep per person total
avg_sleep_graph <- avg_sleep%>%
  ggplot(aes(x=Participant, y = AverageSleep)) +
  geom_bar(stat = "identity", aes(fill=Participant)) +
  geom_text(aes(label=round(AverageSleep,2))) 
print(avg_sleep_graph)

ggplotly(avg_sleep_graph)
  
```





# Calculating total average sleep per person on Day Offs VS No Day Offs
```{r}
# Average sleep per person on Day Offs and No Day Offs
avg_sleep_dayoff_yes_no <- sleep_data %>%
  group_by(Participant,`Day Off`) %>%
  summarise(AverageSleep = mean(`Sleep Duration`))
print(avg_sleep_dayoff_yes_no)


# Graphing - Average sleep per person on Day Offs and No Day Offs
avg_sleep_dayoff_yes_no_graph <- sleep_data %>%
  group_by(Participant,`Day Off`) %>%
  summarise(AverageSleep = mean(`Sleep Duration`)) %>% 
  ggplot(aes(x=Participant, y=AverageSleep, fill = `Day Off`)) +
  geom_bar(stat = "identity", position=position_dodge(width=0.9)) +
  geom_text(aes(label=round(AverageSleep,2)),position=position_dodge(width=0.9)) +
  scale_x_discrete(labels= c("Bishesta"="Bishesta (16 Yrs)","Prashansa"="Prashansa (8 Yrs)", "Rashmina"="Rashmina (43 Yrs)", "Sashi"="Sashi (42 Yrs)")) +
  labs(x="Participants",
       y="Sleep Duration in Hrs")

print(avg_sleep_dayoff_yes_no_graph)

ggplotly(avg_sleep_dayoff_yes_no_graph)

# Add another graph with mutated column

```



# Sleep data of Sashi
```{r}
# All Sleep Data for Sashi
sleep_data_Sashi <- sleep_data %>%
  filter(Participant == "Sashi") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`) 
print(sleep_data_Sashi)

# Sleeping pattern graph
sleep_data_Sashi_graph <- sleep_data_Sashi %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(color="purple") +
  geom_point(aes(color=`Day Off`)) +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Sashi$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Sashi")
ggplotly(sleep_data_Sashi_graph)


```

```{r}
# Comparing amount of sleep during work day and non-working days
sleep_data_Sashi %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(aes(group=`Day Off`, color=`Day Off`))+
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Sashi$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Sashi")
```

```{r}
# Sleep Duration on Day Offs
sleep_data_Sashi_day_off <- sleep_data_Sashi %>%
  filter(Participant == "Sashi" & `Day Off`=="Yes") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Day Offs - Graph
sleep_data_Sashi_day_off %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="blue") +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Sashi$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))


```

```{r}
# Sleep Duration on Work Day
sleep_data_Sashi_day_off_no <- sleep_data_Sashi %>%
  filter(Participant == "Sashi" & `Day Off`=="No") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Work Day - Graph
sleep_data_Sashi_day_off_no %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="red") +
  ylim(c(4,12))  +
  scale_x_continuous(breaks=sleep_data_Sashi$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))

```


# Sleep Data for Rashmina

```{r}
# All Sleep Data for Rashmina
sleep_data_Rashmina <- sleep_data %>%
  filter(Participant == "Rashmina") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`) 
print(sleep_data_Rashmina)

# Sleeping pattern graph
sleep_data_Rashmina_graph <- sleep_data_Rashmina %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(color="purple") +
  geom_point(aes(color=`Day Off`)) +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Rashmina$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Rashmina")
ggplotly(sleep_data_Rashmina_graph)
```

```{r}
# Comparing amount of sleep during work day and non-working days
sleep_data_Rashmina %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(aes(group=`Day Off`, color=`Day Off`))+
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Rashmina$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Rashmina")
```


```{r}
# Sleep Duration on Day Offs
sleep_data_Rashmina_day_off <- sleep_data_Rashmina %>%
  filter(Participant == "Rashmina" & `Day Off`=="Yes") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Day Offs - Graph
sleep_data_Rashmina_day_off %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="blue") +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Rashmina$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))


```

```{r}
# Sleep Duration on Work Day
sleep_data_Rashmina_day_off_no <- sleep_data_Rashmina %>%
  filter(Participant == "Rashmina" & `Day Off`=="No") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Work Day - Graph
sleep_data_Rashmina_day_off_no %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="red") +
  ylim(c(4,12))  +
  scale_x_continuous(breaks=sleep_data_Rashmina$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))

```

# Sleep data for Bishesta

```{r}
# All Sleep Data for Bishesta
sleep_data_Bishesta <- sleep_data %>%
  filter(Participant == "Bishesta") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`) 
print(sleep_data_Bishesta)

# Sleeping pattern graph
sleep_data_Bishesta_graph <- sleep_data_Bishesta %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(color="purple") +
  geom_point(aes(color=`Day Off`)) +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Bishesta$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Bishesta")
ggplotly(sleep_data_Bishesta_graph)
```

```{r}
# Comparing amount of sleep during work day and non-working days
sleep_data_Bishesta %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(aes(group=`Day Off`, color=`Day Off`))+
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Bishesta$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Bishesta")
```


```{r}
# Sleep Duration on Day Offs
sleep_data_Bishesta_day_off <- sleep_data_Bishesta %>%
  filter(Participant == "Bishesta" & `Day Off`=="Yes") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Day Offs - Graph
sleep_data_Bishesta_day_off %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="blue") +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Bishesta$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))


```

```{r}
# Sleep Duration on Work Day
sleep_data_Bishesta_day_off_no <- sleep_data_Bishesta %>%
  filter(Participant == "Bishesta" & `Day Off`=="No") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Work Day - Graph
sleep_data_Bishesta_day_off_no %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="red") +
  ylim(c(4,12))  +
  scale_x_continuous(breaks=sleep_data_Bishesta$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))

```

# Sleep data for Prashansa


```{r}
# All Sleep Data for Prashansa
sleep_data_Prashansa <- sleep_data %>%
  filter(Participant == "Prashansa") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`) 
print(sleep_data_Prashansa)

# Sleeping pattern graph
sleep_data_Prashansa_graph <- sleep_data_Prashansa %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(color="purple") +
  geom_point(aes(color=`Day Off`)) +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Prashansa$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(title="Prashansa")
ggplotly(sleep_data_Prashansa_graph)
```

```{r}
# Comparing amount of sleep during work day and non-working days
sleep_data_Prashansa %>%
  ggplot(aes(x=Date, y=`Sleep Duration`)) +
  geom_line(aes(group=`Day Off`, color=`Day Off`))+
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Prashansa$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Prashansa")
```


```{r}
# Sleep Duration on Day Offs
sleep_data_Prashansa_day_off <- sleep_data_Prashansa %>%
  filter(Participant == "Prashansa" & `Day Off`=="Yes") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Day Offs - Graph
sleep_data_Prashansa_day_off %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="blue") +
  ylim(c(4,12)) +
  scale_x_continuous(breaks=sleep_data_Prashansa$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))


```

```{r}
# Sleep Duration on Work Day
sleep_data_Prashansa_day_off_no <- sleep_data_Prashansa %>%
  filter(Participant == "Prashansa" & `Day Off`=="No") %>%
  select(Participant,Date,Day,`Day Off`,`Sleep Duration`)
# Sleep Duration on Work Day - Graph
sleep_data_Prashansa_day_off_no %>%
  ggplot(aes(x=Date, y=`Sleep Duration`))+
  geom_line(color="red") +
  geom_point(color="red") +
  ylim(c(4,12))  +
  scale_x_continuous(breaks=sleep_data_Prashansa$Date) +
  theme(axis.text.x = element_text(angle=45, hjust=1))

```


# T-test
```{r}
# T-test for Sashi
t.test(sleep_data_Sashi$`Sleep Duration`, mu = 7, alternative = "less")
# T-test for Rashmina
t.test(sleep_data_Rashmina$`Sleep Duration`, mu = 7, alternative = "less")
# T-test for Bishesta
t.test(sleep_data_Bishesta$`Sleep Duration`, mu = 8, alternative = "less")
# T-test for Rashmina
t.test(sleep_data_Prashansa$`Sleep Duration`, mu = 9, alternative = "less")
```


# One sample proportion test

```{r}
prop.test(x=0, n=4, p=0.5, alternative="greater")
```




